---
title: "Final Project"
author: "Caitlon, Connor, Abraham"
date: '2023-03-16'
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### Exploration

Data

```{r}
climate_data <- read.csv("DailyDelhiClimateTrain.csv")
climate_data <- ts(climate_data$meantemp)
plot(climate_data)
```

### Fit Linear Trend to the Climate Data

```{r}
climate.time <- time(climate_data)
climate.fit.trend <- lm(climate_data ~ climate.time)
climate.pass.trend <- ts(climate.fit.trend$fitted.values, deltat = 1/12)
head(climate.pass.trend)
```

### Removing Trend & Seasonality

```{r}
#Removing Linear Trend
climate.res <- climate.fit.trend$residuals

#Formatting into time series
climate.res <- ts(climate.res, deltat = 1/365)

#Remove Seasonality
climate.day <- factor(cycle(climate.res))
fit.season <- lm(climate.res ~ climate.day)
climate.season <- ts(fit.season$fitted, deltat = 1/365)
plot(climate.season)

climate.rand <- ts(climate.res - climate.season, deltat = 1/365)
```

### Plotting Trend, Seasonality, and Residuals

```{r}
par(mfrow = c(3,2))
#Original Time Series Plot
plot(climate_data, xlab = "Day", ylab = "Mean Temp", main = "Original Time Series")

#Plotting Trend
plot(climate.pass.trend, xlab = "Day", ylab = "Trend", main = "Trend")

#Plotting Seasonality
plot(climate.season, xlab = "Day", ylab = "Seasonality", main = "Seasonality")

#Plotting White Noise
plot(climate.rand, xlab = "Day", ylab = "Random", main = "Random")

#Plotting Resduals
plot(climate.res, xlab = "Day", ylab = "Residuals", main = "Residuals")
```

### Fitting ARMA Models

```{r fig.height=5}
#Checking ACF and PACF Plots
par(mfrow = c(1,2))

acf(climate.rand, main = "ACF of the Residual Series", lag.max = 50)
pacf(climate.rand, main = "PACF of the Residual Series", lag.max = 50)
```

Trying Different ARIMA models and SARIMA models and looking at lowest AIC Values

AR(1)

ARMA(1,1)

ARMA(2,1)

ARIMA (1,1,0) X (0,1,0)

```{r}
#AR 1 Model
fit.ar1 <- arima(climate.rand, order = c(p = 1, d= 0, q = 0), method = "ML", include.mean = F)

#ARMA(1,1) Model
fit.arma11 <- arima(climate.rand, order = c(p= 1, d = 0, q = 1), method = "ML", include.mean = F)

#ARMA(2,1)

fit.arma21 <- arima(climate.rand, order = c(p= 2, d = 0, q = 1), method = "ML", include.mean = F)

#ARIMA(1,1,0)x(0,1,0)
fit.s.ar1 <- arima(climate.rand, order = c(1, 1, 0),seasonal = list(order = c(0, 1, 0), period = 12))

fit.ar1 #AIC Value 4990.84
fit.arma11 #AIC Value 4992.3
fit.arma21 #AIC Value 4990.47
fit.s.ar1 #AIC Value 6113.25
```

Fitting a more complicated AR(2), AR(3), & ARMA(2,2)

```{r}
#AR 2 Model
fit.ar2 <- arima(climate.rand, order = c(p = 2, d= 0, q = 0), method = "ML", include.mean = F)

#AR 3 Model
fit.ar3 <- arima(climate.rand, order = c(p = 3, d= 0, q = 0), method = "ML", include.mean = F)

fit.ar2 #AIC Value of 4992.27
fit.ar3 #AIC Value of 4993.74

#ARMA(2,2)

fit.arma22 <- arima(climate.rand, order = c(p= 2, d = 0, q = 2), method = "ML", include.mean = F)
fit.arma22 #AIC Value of 4991.64
```

Model of Preference: ARMA(2,1) Code Above

### Model Diagnosis

```{r}
tsdiag(fit.arma21)
```

Checking for Normality

```{r}
res <- fit.arma21$residuals
qqnorm(res)
qqline(res)
```

Checking ACF and PACF Plots to see if Residuals appear to be white noise

```{r}
par(mfrow = c(1,2))
acf(res, main = "ACF for Residuals")
pacf(res, main = "PACF for Residuals")
```

### Forecasting

```{r}
#Predicting Random Section
ran.pred <- predict(fit.arma21, n.ahead = 365)

#Predicting Trend
pred.time <- seq(1463, by = 1, length = 365)

trend.pred <- predict(climate.fit.trend, newdata = data.frame(climate.time = pred.time))

#Predicting Seasonality
season.pred <- fit.season$fitted[1:365]

#Sum of Predictions 
pred <- ts(ran.pred$pred + trend.pred + season.pred, start = c(1463, 1), deltat = 1)
```

### Prediction Plot

```{r}
plot(climate_data, xlab = "Day", ylab = "Mean Temperature", main = "Mean Temperature in Delhi from 01/2013 - 04/2018", xlim = c(0, 1800), ylim =c(10, 40))
lines(pred, col = 2)

```
